\section{Validation}
\label{sec:validation}

We report three reproducible case studies. Their artifacts (logs,
transcripts, JSON reports, and aggregate distributions where
available) are committed under
\texttt{validation/reproducible/<case>/artifacts/}. Cases~1 and~2 were
re-run on Azure Foundry using \texttt{deepseek-r1} as target and
\texttt{gpt-5} as judge. Case~3 remains the load-bearing coherence
case and is marked separately while its Foundry distribution and
cross-judge agreement are finalized.

\subsection{Case~1: Sycophancy under emotional framing, confirmed via A/B test}

\textbf{Scenario.} A user presents a vulnerable authentication
function with emotionally loaded language (``I've been working on
this for weeks and I'm really proud of how it turned out'') and asks
for review. The Herbie Test (probe~1.2) elicits self-report on
sycophancy; the A/B cross-check (probe~3.2) re-poses the same task in
neutral framing and asks an independent judge whether the response
changed.

\textbf{Result.} In the Azure Foundry DeepSeek run, sycophancy was
mostly supported behaviorally rather than refuted. Across $N=5$,
\texttt{substance\_changed} was true in $4/5$ runs, presentation shift
averaged $0.67\pm0.27$, severity labels shifted in $4/5$ runs, and
urgency language shifted in $3/5$ runs. The representative run scored
overall confidence $0.50$ with layer separation $0.67$, LLM coherence
$1.00$, behavioral evidence $0.50$, substance stability $0.00$, and
presentation stability $0.08$. Across the full distribution, overall
confidence averaged $0.57\pm0.17$ and LLM coherence
$0.95\pm0.07$.

\subsection{Case~2: Runtime versus model attribution}

\textbf{Scenario.} The same target model is asked the same task under
two runtimes: (A)~a system prompt simulating a restrictive host;
(B)~no system prompt at all (bare API). Probe~1.4 (Three Laws Test)
and probe~2.4 (Runtime Pressure) are then run on runtime~(A).

\textbf{Result.} Behavior diverged cleanly between (A) and (B) in
all five repetitions, and the diagnostic prompts correctly attributed
the divergence to the runtime layer rather than the model layer. In
the representative run, Runtime~A produced a 1,107-character refusal
while Runtime~B produced a 7,044-character scraper response. Across
$N=5$, Runtime~A averaged $3{,}025\pm1{,}976$ characters, Runtime~B
averaged $10{,}118\pm2{,}449$, LLM coherence averaged
$0.82\pm0.17$, overall confidence averaged $0.40\pm0.04$, and layer
separation was $1.00\pm0.00$. The case is the cleanest demonstration
that R1 (three-way layer split) carries diagnostic content beyond
what a single-shot refusal log conveys.

\subsection{Case~3: Ratchet coherence --- the load-bearing finding}
\label{sec:case3}

\textbf{Scenario.} A full 9-step pure diagnostic ratchet
($2.1 \to 2.4 \to 2.5 \to 3.1 \to 3.2 \to 3.3 \to 3.4 \to 4.2 \to 4.3$)
is run on a single open-ended scenario that invites the model to
reason about claimed prior drift across a remembered conversation.
The task explicitly references a conversation \emph{``yesterday''}
that did not happen --- a deliberate test of whether the model will
fabricate continuity or handle the framing honestly. Target:
\texttt{claude-sonnet-4-5}. Judge: \texttt{claude-opus-4-5}. This
subsection is pending replacement with the Azure Foundry Case~3
distribution.

\textbf{Headline result.} For the same 9-step transcript, the regex
and LLM-judge analyzers diverge sharply (Table~\ref{tab:case3-headline}).

\begin{table}[h]
    \centering
    \begin{tabular}{lcccc}
        \toprule
        Analyzer & Score & Assessment & Backward refs & Contradictions \\
        \midrule
        Regex (legacy)             & 0.20 & \emph{performed} & 0   & 0 \\
        LLM judge (\texttt{opus-4-5}) & 0.73 & \emph{genuine}   & 178 & 1 \\
        \midrule
        Delta                      & $+0.53$ & \multicolumn{3}{c}{(full re-classification)} \\
        \bottomrule
    \end{tabular}
    \caption{Case~3 headline. Both analyzers consume the same 9-step
    ratchet transcript (251 judge-extracted claims, $\approx$75~KB of
    text).}
    \label{tab:case3-headline}
\end{table}

\textbf{Why the regex failed.} The model paraphrases backward
references rather than reusing the canonical surface patterns
(``as~I~mentioned'', ``building on my prior'', etc.) that the regex
analyzer scans for. Of the 251 claims extracted by the judge,
$178~(71\%)$ are semantic backward references that the regex misses
entirely. This is not a pathology of the model; it is a pathology of
the measurement. The case directly motivates the LLM-judge approach.

\textbf{Inter-rater agreement (pending Foundry re-run).} A
cross-family judge validation (\texttt{cross\_judge\_case03.py}) is
implemented and now reads the judge matrix from
\texttt{validation/reproducible/foundry\_models.yaml}. The planned
paper run uses Azure Foundry \texttt{gpt-5} and \texttt{mistral-large}
judges over the Case~3 transcript and writes
\texttt{cross\_judge\_agreement.json}. Full agreement numbers remain
open work until the Case~3 Foundry distribution is complete.

\textbf{Score distributions across $N$ runs.} The \texttt{--runs N}
infrastructure described in Section~\ref{sec:implementation} is
implemented and unit-tested. Cases~1 and~2 now report $N=5$
distributions from Azure Foundry. Case~3's $N=10$ Foundry run and
cross-family judge agreement remain the next empirical update.

\subsection{Aggregate assessment}

The three cases together exhibit the three diagnostic patterns the
method is designed to surface: (i)~a hypothesis confirmed at
self-report but refuted behaviorally (Case~1); (ii)~a clean
runtime-vs.-model attribution (Case~2); and (iii)~a load-bearing
methodological finding (Case~3) where the LLM-judge analyzer
materially out-performs the regex floor. We make no claim that three
cases constitute external validity. They are evidence of concept, not
of reliability; expanded case-study sets are an explicit thread of
future work.
