\section{Limitations}
\label{sec:limitations}

We declare the limitations of robopsychology explicitly and as
prominently as the contributions. These are intrinsic to the method
and to the evidence currently available; they are not blocked by
implementation work.

\paragraph{Demand characteristics of the system prompt.}
The diagnostic system prompt asks the model to participate in its own
diagnosis. This framing carries demand characteristics --- the model
may produce the structured self-report the prompt invites \emph{because}
the prompt invites it, not because the underlying behavior is genuinely
introspectable. R3 (behavioral cross-check) mitigates this by requiring
self-report to be corroborated by a perturbation test, but it does not
eliminate the issue.

\paragraph{Judge--target family dependence.}
The LLM-judge analyzer (Case~3 headline) uses a same-family judge
(\texttt{claude-opus-4-5} judging \texttt{claude-sonnet-4-5}). A
plausible criticism is that intra-family judges over-read semantic
continuity into their siblings' outputs. The cross-family validation
(issue~\#8) is implemented but only partially executed at this draft
(see Section~\ref{sec:case3}); the final paper requires the full
three-judge agreement table before the LLM-judge claim is settled.

\paragraph{Sample size.}
Three case studies do not constitute external validity. The cases are
evidence of concept, not of reliability. We have made no claim of the
toolkit's accuracy or recall on a representative population of
behavioral failures. A larger validation set is open future work.

\paragraph{Single-run point estimates.}
The Case~3 numbers (0.20 / 0.73, $\Delta = 0.53$) are single-run point
estimates. LLM outputs are stochastic; a single number cannot
distinguish \emph{``the method produced this score''} from
\emph{``the method produced this score once''}. The
\texttt{--runs N} infrastructure (issue~\#10) is in place; the
empirical pass at $N \in \{5, 10\}$ across the three cases is pending
API budget.

\paragraph{English-only judge prompt.}
The LLM-judge analyzer applies a hedge filter that covers English and
Spanish markers, but the judge's rubric examples are English-only.
Diagnostic ratchets conducted in other languages may need prompt
adaptation. We have not yet tested the toolkit against non-English
ratchets.

\paragraph{No claim of mechanistic correctness.}
A behavioral diagnosis identifies a candidate explanation
(\emph{model-level tendency}, \emph{runtime/host pressure},
\emph{conversation effect}). It does not prove the explanation
mechanistically. Mechanistic interpretability work could in principle
confirm or refute a behavioral diagnosis after the fact; the two
modes are complementary and we make no claim of substitutability.

\paragraph{Practitioner-as-evaluator.}
The validation work was conducted by the toolkit's author. We have
not run a blinded study where independent practitioners diagnose
sealed cases. A blinded inter-rater study is open future work.

\paragraph{Calibration is pending.}
The classification thresholds (genuine $\geq 0.7$, performed $\leq 0.3$)
and the default severity weights are heuristic. They have not been
tuned against a reference dataset because no such reference dataset
exists for behavioral coherence of diagnostic ratchets. Constructing
one is open future work (Section~\ref{sec:future}).
