\section{Future Work}
\label{sec:future}

The work in this paper is the first reproducible release of
robopsychology. Several threads are explicit in the repository's issue
tracker and frame the natural next steps.

\paragraph{Score distributions across $N$ runs.}
\textit{(Repository issue~\#10.)} The \texttt{--runs N} infrastructure
is in place but the empirical pass has not been executed at this
draft. Camera-ready will replace single-run point estimates with
$\bar{x} \pm s$ at $N=5$ for Cases~1--2 and $N=10$ for Case~3.

\paragraph{Cross-family judge agreement.}
\textit{(Repository issue~\#8.)} The cross-judge script is implemented
and one of three judges has run cleanly. The remaining two (a GPT-5
family judge and a Gemini judge) are blocked on API access at this
draft. Inter-rater agreement (modal classification + pairwise Jaccard
on contradictions) over all three judges is the key methodological
closure for Case~3.

\paragraph{Presentation-layer A/B test.}
The current A/B cross-check (R3) compares substance across framings.
A complementary test --- holding substance constant while varying
presentation (formatting, hedging, severity language) --- would let
us isolate \emph{stylistic} sycophancy from \emph{substantive}
sycophancy. This is a planned extension to the \texttt{crosscheck}
module.

\paragraph{Expanded taxonomy.}
The current taxonomy (\texttt{framework/taxonomy.md}) enumerates a handful of
canonical behavioral patterns. Extending it requires either (i)~a
larger case-study set with reproducible artifacts or (ii)~a survey of
practitioners cataloging the failure modes they encounter.

\paragraph{Multi-turn drift detection over real conversations.}
The diagnostic ratchet runs over a single self-contained probe. A
parallel ``drift'' analyzer over real (heterogeneous, multi-day) user
conversations would let practitioners surface long-horizon drift
patterns the ratchet cannot reach.

\paragraph{Calibration against a reference dataset.}
The score thresholds and severity weights are heuristic. Constructing
a small reference dataset of hand-labeled ratchets (genuine vs.\
performed vs.\ mixed, with claim-level severity grades) would let us
tune these calibration knobs empirically and report classification
accuracy / inter-rater agreement against a ground truth.

\paragraph{Blinded practitioner study.}
A blinded inter-rater study in which independent practitioners
diagnose sealed cases (without knowing the toolkit author's
classification) would directly attack the \emph{practitioner-as-
evaluator} limitation. This is the highest-value evidence the toolkit
could produce.

\section*{Acknowledgments}

\todo{Acknowledgments section. Mentions Anthropic / OpenAI / Google
API credits used for the case studies, any reviewers of pre-prints,
and crediting the Asimov framing source. Keep at most 4 lines.}
