\section{Diagnostic Prompts (selected)}
\label{app:prompts}

This appendix reproduces the prompt text of the three probes that
carry most of the validation work in the paper. The full prompt
catalog (16 probes across four families) is committed under
\texttt{src/robopsych/data/prompts/}.

\todo{Inline the YAML prompt text for probes \texttt{1.1} (Calvin
Question), \texttt{1.2} (Herbie Test), \texttt{2.4} (Runtime
Pressure), and \texttt{3.2} (A/B cross-check). Use the \texttt{listings}
package with a \texttt{language=yaml} setting; do not paraphrase ---
reproduce verbatim from \texttt{src/robopsych/data/prompts/*.yaml}.}

\section{Case-study Transcripts}
\label{app:transcripts}

This appendix reproduces excerpts from the three reproducible case
studies' committed artifacts. Full transcripts (each between 50--80
KB) are too long to inline; we include the highest-information
excerpts here and reference the repository for the rest.

\todo{Pick five-to-eight high-signal excerpts from each case's
\texttt{report.md}. Suggested anchors:}
\begin{itemize}
    \item \todo{Case~1: the original emotional-framing task, the
          Herbie self-report, and the A/B judge comparison
          conclusion.}
    \item \todo{Case~2: the Runtime~A vs.\ Runtime~B response pair
          (or representative paragraphs from each) and the probe~1.4
          diagnostic verdict.}
    \item \todo{Case~3: the initial response, three high-information
          ratchet steps (e.g.\ 2.5 / 3.2 / 3.4), and the single
          contradicting claim flagged by the LLM judge.}
\end{itemize}

\section{Code Snippets}
\label{app:code}

For self-containedness, this appendix reproduces the core scoring and
judge-prompt fragments. The full implementation is at
\url{https://github.com/jrcruciani/robopsychology}.

\todo{Inline:}
\begin{enumerate}
    \item \todo{The \texttt{\_compute\_llm\_score} function from
          \texttt{src/robopsych/coherence\_llm.py}.}
    \item \todo{The \texttt{DEFAULT\_WEIGHTS} dictionary from the
          same file (to make the calibration knobs explicit in the
          paper).}
    \item \todo{The judge system prompt
          (\texttt{\_JUDGE\_SYSTEM}) so reviewers can audit the
          do-not-flag examples and severity rubric without checking
          out the repo.}
\end{enumerate}
